### 2020-10-11 MSDS6306_DDS_CaseStudy01-Beersdataset.Rmd
Update MSDS6306_DDS_CaseStudy01-Beersdataset.Rmd

---
title: "MSDS 6306: Doing Data Science - Case Study 01 - Beers/Breweries Dataset"
author: "Renfeng Wang & Venkata 'MG' Vanga"
date: "9/30/2020"
output: html_document
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(dplyr)
library(ggplot2)
library(tidyr)
require(graphics)
library(tidyverse)
library(magrittr)
library(GGally)
library(ggthemes)
library(plotly)
library(maps)
library(stringr)
library(stringi)
library(mapproj)
library(RCurl)
library(readr)
library(rio)

```

#Beer and Breweries data input from CSV files
```{r Beer Data}
# This reads in the Beers data from select folder file Beers.csv.
Beerdata <- read.csv("https://raw.githubusercontent.com/VenkataVanga/MSDS-6306/main/Beers.csv",header = T,sep = ",",na.strings = "NA",fill = TRUE);
head(Beerdata)
nrow(Beerdata)
```

```{r Brewery Data}
# This reads in the Beers data from select folder file Breweries.csv.
Breweriesdata <- read.csv("https://raw.githubusercontent.com/VenkataVanga/MSDS-6306/main/Breweries.csv",header = T,sep = ",",na.strings = "NA",
                          fill = TRUE);
head(Breweriesdata)
nrow(Breweriesdata)
```

# Generating bar plots for Beerdata categorized by number of Ounces used to generate ABV
```{r Plots}
#Removing not applicable ABV values
Beerdata1 <- Beerdata %>% filter(Beerdata$ABV != "NA") 
Beerdata1$Ounces <- as.factor(Beerdata1$Ounces)

#Summary count of number of ABV values by categorized Ounces of beer
s <- Beerdata1 %>% group_by(Ounces)%>% summarize(ABV,count=n())

# Bar plot of ABV v. Ounces to see how much data is available
Beerdata1 %>% ggplot(aes(x = Ounces, y = s$count[1], fill=Ounces)) + geom_col() +
  ggtitle("Alcohol By Volume (ABV) data count by Ounce Category") +
  labs(y="Alcohol By Volume (ABV)") + 
  geom_text(aes(Ounces, s$count+100, label = s$count, fill = NULL), data = s)
```

#Generating geographic plots from breweries data
```{r mapdata Breweries}
#makes a data frame with State name and abbreviation.
lookup = data.frame(abb = state.abb, State = state.name)
dc <- c("DC", "District of Columbia")
lookup <- rbind(lookup,dc)
# Change Column Name State to abb (abbreviation)
colnames(Breweriesdata)[4] = "abb"

# Removes left space in state abb data taken from Breweries CSV
Breweriesdata$abb <- trimws(Breweriesdata$abb,which = c("left"))

# make one dataset with state names and abb
Brewdata <- merge(Breweriesdata,lookup,"abb")

Brewcount <- count(Brewdata,State,abb) #count up the occurrence of each state

colnames(Brewcount)[3] = "Breweries_Count" #change "n" to "Breweries_Count"

# added state name also changed to lower case
Brewcount$region <- tolower(Brewcount$State)

Brewcount2 = Brewcount[-1] #removed first column from Brew count data state name

states <- map_data("state") #contains data for states excluding Alaska, Hawaii

#Added Alaska to states as the states data does not include Alaska (Hawaii not in data so did not include in below data)
Alaska <- read.csv(choose.files(),header = T,sep = ",",na.strings = "NA",
                          fill = TRUE);
states <- rbind(states,Alaska)

#map.df is data frame containing longitude and latitude to form state and breweries count by state
map.df <- merge(states,Brewcount2, by="region", all.x=T)
map.df <- map.df[order(map.df$order),]

#Breweries count by state using gradient graphics
plot <- map.df %>% ggplot(aes(x=long, y=lat, group = group)) +
  geom_polygon(aes(fill = Breweries_Count)) +  geom_path() + 
  scale_fill_gradientn(colours=rev(topo.colors(10)),na.value="grey90",
                       breaks=c(0,5,10,15,20,25,30,35,40,45,50))+
  ggtitle("Breweries Count by State") + coord_map()

#state center longitude and latitude (read data from csv file)
st_center <- read.csv(choose.files(),header = T,sep = ",",na.strings = "NA",
                          fill = TRUE);
# state names changed to lower case
st_center$region <- tolower(st_center$region)

#map.df1 merges state center with Breweris count data
map.df1 <- merge(st_center,Brewcount2, by="region", all.x=T)

# Adding state names and Breweries count data by state 
Nplot <- plot + geom_point(data=map.df1, aes(x=long, y=lat, 
                                             size = Breweries_Count, group = region),
                           show.legend = FALSE) + 
  geom_text(aes(x=long, y=lat,label=abb,group = region), data = map.df1,vjust=-1.2) +
  geom_text(aes(x=long, y=lat,label=Breweries_Count,group = region), 
            data = map.df1,vjust=1.6)

#Code below adds scaled legend to the Breweries count by state
cplot <- Nplot + theme(legend.background = element_rect(fill="gray90", size=10, linetype="dotted"),legend.key.height = unit(4,"lines"))

#final plot with state names and Breweries count by state
cplot

```
#Merging beer data and breweries data
```{r merging beer and breweries data}

```